---
title: "From pamDP to DwC"
author: "Marius Bottin"
format: 
  gfm:
    df-print: kable
editor: visual
---

This document is to show quickly how to analyse the structure and variables of pamDP data formats and to transform it to the DarwinCore format (event + Audiovisual Media Description + registers).
Of course, it would be better to make a python code that could be integrated with the codes the soundscape team develop in the institute, but the objective here is only to show a proof of concept.

# Reading data
I put the data in the a directory `data_transformacion` at level -1 of this repository.

First, we read the mapping xlsx file:

```{r}
require(openxlsx)
wb<-loadWorkbook("../data_transformacion/Sistematización de datos - pamDP.xlsx")
mapping<-lapply(names(wb),function(x)read.xlsx(xlsxFile=wb,sheet=x))
names(mapping)<-names(wb)
```

The structure of the mapping file is somewhat dirty, if we want to automatize more the process in the future, we will need to clean that:

```{r}
lapply(mapping,colnames)
```


Now we read the csv files containing the data:

```{r}
csvFiles<-c(media="../data_transformacion/media.csv",deployments="../data_transformacion/deployments.csv",observations="../data_transformacion/observations.csv")
data<-lapply(csvFiles,read.csv)
```

# Checking on variables

The name of the fields are not very clean neither:

```{r}
mapping$media.csv$pamDP.Field.Name
mapping$deployments.csv$`pamDP.-.Field.Name`
mapping$observation.csv$`pamDP.-.Field.Name`
```

We will need to trim the names of the fields (to suppress the ' *') at the end of the character strings:

```{r}
(mapping$media.csv$pamDP_field<-gsub(" \\*$","",mapping$media.csv$pamDP.Field.Name))
(mapping$deployments.csv$pamDP_field<-gsub(" \\*$","",mapping$deployments.csv$`pamDP.-.Field.Name`))
(mapping$observation.csv$pamDP_field<-gsub(" \\*$","",mapping$observation.csv$`pamDP.-.Field.Name`))
```

Now we can check which fields from the mappings are present in data files

```{r}
(m_media<-match(colnames(data$media),mapping$media.csv$pamDP_field))
(m_deployments<-match(colnames(data$deployments),mapping$deployments.csv$pamDP_field))
(m_observations<-match(colnames(data$observations),mapping$observation.csv$pamDP_field))

```

We put in the mapping whether the variables are in the data or not:

```{r}
mapping$media.csv$presentInData<-F
mapping$media.csv$presentInData[m_media]<-T
mapping$deployments.csv$presentInData<-F
mapping$deployments.csv$presentInData[m_deployments]<-T
mapping$observation.csv$presentInData<-F
mapping$observation.csv$presentInData[m_observations]<-T
```

Let's consider that the required variables for DwC are the ones marked as required for pamDP + the ones not having any equivalent in pamDP, that were added in the mapping. 
In the future, we should have a system to map from DwC to all the variables allowing to create the DwC variables, and to document whether the variable is required in the DwC.

```{r}
mapping$media.csv$DwC_required<-ifelse(is.na(mapping$media.csv$Required),T,mapping$media.csv$Required)
mapping$media.csv$DwC_required[is.na(mapping$media.csv$DarwinCore)]<-F
mapping$deployments.csv$DwC_required<-ifelse(is.na(mapping$deployments.csv$Required),T,mapping$deployments.csv$Required)
mapping$deployments.csv$DwC_required[is.na(mapping$deployments.csv$DarwinCore)]<-F
mapping$observation.csv$DwC_required<-ifelse(is.na(mapping$observation.csv$Required),T,mapping$observation.csv$Required)
mapping$observation.csv$DwC_required[is.na(mapping$observation.csv$DarwinCore)]<-F
```

Tables present vs required:

```{r}
table(mapping$media.csv[c("DwC_required","presentInData")])
table(mapping$deployments.csv[c("DwC_required","presentInData")])
table(mapping$observation.csv[c("DwC_required","presentInData")])
```

## Missing variables

Media.csv:

```{r}
mapping$media.csv[!mapping$media.csv$presentInData&mapping$media.csv$DwC_required,c("pamDP.Field.Name","DarwinCore")]
```

Deployments.csv

```{r}
mapping$deployments.csv[!mapping$deployments.csv$presentInData&mapping$deployments.csv$DwC_required,c("pamDP.-.Field.Name","DarwinCore")]
```

Observation.csv

```{r}
mapping$observation.csv[!mapping$observation.csv$presentInData&mapping$observation.csv$DwC_required,c("pamDP.-.Field.Name","DarwinCore")]
```

Para ir más allá y hacer el trabajo de manera más fina, deberíamos
